---
title: "Stats105_Hw2"
author: "Yuhshang"
date: "1/21/2020"
output: pdf_document
---


# Problem 1

## (1)
Note that
$$x_{0.975}-\mu = \sigma *Z_{0.975}$$
Where need to satisfy $\mu \sim N(\mu_{0},\tau_{0})$
$$\tau = 1/\sigma^2$$
$$Z_{0.975}=qnorm(0.975,1)$$
```{r}
mu0  = 40

sig0 = (42-mu0) / 1.96
print(sig0^2)
tau0 = 1/sig0^2
print(tau0)
```

conjugate Normal prior $\mu \sim N(40,\tau_{0}=0.9603647)$ 

## (2)

```{r}
lst = c(38.6,42.4,57.5,40.5,51.7,67.1,33.4,60.9,64.1,40.1,40.7,6.4)
mu1 = sum(lst)/length(lst)
tau1 = 1/100
print(mu1)
print(tau1)
```
$$\mu_{1} \sim N(45.28333,0.01)$$
Since nomal prior, normal likelihood gives conjugate distribution.

Then, posterior is 
$$ N(\frac{\tau_{0}}{n\tau_{1}+\tau_{0}}\mu_{0}+\frac{n\tau_{1}}{n\tau_{1}+\tau_{0}}\mu_{1},\frac{1}{n\tau_{1}+\tau_{0}})$$

```{r}
n   = length(lst)
mu3 = tau0*mu0/(n*tau1+tau0) + n*tau1*mu1/(n*tau1+tau0) 
std2 = 1/(n*tau1+tau0) 
print(n)
print(mu3)
```

$$N_{post}(40.58684,0.9256134)$$
The posterior's mean also follows a normal distriution, the relative contribution is to increase the mean by a bit and reduces its std by a little bit. 

## (3)

```{r}
x_lst = seq(30,55,by=0.001)
plot(x_lst,dnorm(x_lst,mu0,(1/tau0)^0.5),type='l',col='blue',xlab='x',ylab='probability',ylim=c(0,.5),lwd = 2.5)
lines(x_lst,dnorm(x_lst,mu3,(std2)^0.5),type='l',col='red',xlab='x',ylab='probability',ylim=c(0,.5),lwd = 2.5)
legend('topleft',col=c('blue','red'),legend= c('prior distribution','posterior distribution'),lwd =c(2.5,2.5),cex = 0.65)
```
The posterior mean larger than that of pior. The new distribution is like a shift of prior to right and the mean gets a bit larger.


## (4)
```{r}
print(qnorm(c(0.1,0.9),mu3,(std2)^0.5))
```

$$P_{post}(\mu<39.35387|Data)=0.1$$
$$P_{post}(\mu>41.81976|Data)=0.9$$

## (5)
Note that $log \mu$ has mo definition if $\mu \leq 0$
and that since $p(\mu|x) \sim e^{-x^2}$ which converges much faster that $\log \mu$ when x gets larger. Thus $log \mu P(\mu|x)->0$ as $x-> infinity$

Thus, 
$$E[\log \mu] = \int_{0^+}^{Inf} \log \mu \cdot p(\mu|x) dx \approx \int_{qnorm(0.0000001,mu_{post},1/tau_{post})}^{qnorm(0.9999999,mu_{post},1/tau_{post})} \log \mu \cdot p(\mu|x) dx \approx 3.703161$$
```{r}
interva = qnorm(c(0.0000001,0.9999999),mu3,std2^0.5)
fun <- function(x) log(x) * dnorm(x,mu3,std2^0.5)
Q <- integrate(fun,interva[1],interva[2])
print(Q)
    
```


# Problem 2
## (1)

$$\lambda \sim Gamma(a,b)$$
$$Y|\lambda \sim poisson (N\lambda)$$

$$\lambda_{posterior}|Y \sim Gamma(a+Y,b+N)$$
Liklihood,
$$f(Y|\lambda)=\frac{e^{-N\lambda}\cdot(N\lambda)^y}{y!}$$
Pior,
$$f(\lambda) = \frac{b^a}{\gamma(a)}\lambda^{a-1}e^{-b \lambda}$$
$$f(\lambda|y)=\frac{f(Y|\lambda)f(\lambda)}{f(y)}$$
$$f(\lambda|y) \propto e^{-\lambda N}(N \lambda)^y \cdot  \lambda^{a-1}e^{-b \lambda} \propto \lambda^{y+a-1}e^{-(\lambda+b)\lambda}$$
$$\lambda \sim Gamma(y+a,N+b)$$
Assum2 $Y_{i} \sim i.i.d Poisson(N \lambda)$
$$f( Y_{1},Y_{2}...Y_{n}|\lambda)= f( Y_{1}|\lambda) \cdot f( Y_{2}|\lambda)\cdots f( Y_{n}|\lambda)$$

$$f( \lambda|Y_{1},Y_{2}...Y_{n}) = \frac{f( Y_{1}|\lambda) \cdot f( Y_{2}|\lambda)\cdots f( Y_{n}|\lambda) f(\lambda)}{f(Y_{1},Y_{2}...Y_{n})}$$

$$f( \lambda|Y_{1},Y_{2}...Y_{n})\propto e^{-\lambda N}(N \lambda)^{y_{1}} \cdot e^{-\lambda N}(N \lambda)^{y_{2}} \cdots e^{-\lambda N}(N \lambda)^{y_{n}}\cdot  \lambda^{a-1}e^{-b \lambda} $$
$$ \propto e^{-(nN+b) \lambda}(N \lambda)^{a+\sum_{i}^{n} y_{i}}$$

$$\lambda_{posterior}|Y \sim Gamma(a+\sum_{i}^{n}Y_{i},b+nN)$$

```{r}
Y_lst = c(rep(1,12),rep(2,6),rep(10,2))
print(sum(Y_lst))
```
## (2)

$$\lambda_{posterior}|Y \sim Gamma(44.01,50.01)$$
```{r}
a = 0.01
b = 0.01
N = 50
ap =  sum(Y_lst)+a
bp =  (N+b)
x_lst = seq(0,10,by=0.001)
y_lst = dgamma(x_lst,shape=ap,scale=1/bp)
plot(x_lst,y_lst,type='l',col='blue',xlab='lambda',ylab='probability',lwd = 2.5,main='posterior')
inter = qgamma(c(0.025,0.975), shape=ap, scale =1/bp)
I= x_lst>qgamma(.025,shape=ap,scale=1/bp)
x_lstnew = x_lst[I==TRUE]
I= x_lstnew <qgamma(.975,shape=ap,scale=1/bp)
y_lstnew = dgamma(x_lstnew,shape=ap,scale=1/bp)
polygon(c(x_lstnew[I],rev(x_lstnew[I])),c(y_lstnew[I],
+ rep(0,sum(I))),col="red",density=30,border=NA)
abline(v = inter[1], col="red", lwd=3, lty=2)
abline(v = inter[2], col="red", lwd=3, lty=2)
```
```{r}
print(inter)
mean1 =  ap/bp
print(mean1)
```

The 95% credible interval is (0.6394519,1.1584123) and mean is 0.880024

```{r}
a = 0.01
b = 0.01
N = 50
ap =  sum(Y_lst)+a
bp =  (N+b)
x_lst = seq(0,10,by=0.001)
y_lst = dgamma(x_lst,shape=ap,scale=1/bp)
plot(x_lst,y_lst,type='l',col='blue',xlab='lambda',ylab='probability',lwd = 2.5)
inter = qgamma(c(0.025,0.975), shape=ap, scale =1/bp)
print(inter)
I= x_lst>qgamma(.025,shape=ap,scale=1/bp)
x_lstnew = x_lst[I==TRUE]
I= x_lstnew <qgamma(.975,shape=ap,scale=1/bp)
y_lstnew = dgamma(x_lstnew,shape=ap,scale=1/bp)
mean1 =  ap/bp
print(mean1)
abline(v = mean1, col="blue", lwd=1.2, lty=2)

a = 1
b = 1
N = 50
ap =  sum(Y_lst)+a
bp =  (N+b)
x_lst = seq(0,10,by=0.001)
y_lst = dgamma(x_lst,shape=ap,scale=1/bp)
lines(x_lst,y_lst,type='l',col='red',xlab='lambda',ylab='probability',lwd = 2.5)
inter2 = qgamma(c(0.025,0.975), shape=ap, scale =1/bp)
print(inter2)
I= x_lst>qgamma(.025,shape=ap,scale=1/bp)
x_lstnew = x_lst[I==TRUE]
I= x_lstnew <qgamma(.975,shape=ap,scale=1/bp)
y_lstnew = dgamma(x_lstnew,shape=ap,scale=1/bp)
mean2 =  ap/bp
abline(v = mean1, col="black", lwd=1.2, lty=2)
print(mean2)

a = 0.1
b = 0.1
N = 50
ap =  sum(Y_lst)+a
bp =  (N+b)
x_lst = seq(0,10,by=0.001)
y_lst = dgamma(x_lst,shape=ap,scale=1/bp)
lines(x_lst,y_lst,type='l',col='yellow',xlab='lambda',ylab='probability',lwd = 2.5)
inter3 = qgamma(c(0.025,0.975), shape=ap, scale =1/bp)
print(inter3)
I= x_lst>qgamma(.025,shape=ap,scale=1/bp)
x_lstnew = x_lst[I==TRUE]
I= x_lstnew <qgamma(.975,shape=ap,scale=1/bp)
y_lstnew = dgamma(x_lstnew,shape=ap,scale=1/bp)
mean3 =  ap/bp
abline(v = mean3, col="black", lwd=1.2, lty=2)
print(mean3)

legend('topright',col=c('blue','red','yellow'),legend= c('pior a =b = 0.01','pior a =b = 1','pior a =b = 0.1'),lwd =c(2.5,2.5),cex = 0.65)
```


```{r}
qtls = matrix(NA,3,3)
qtls[1,] = c(inter[1],inter[2],mean1)
qtls[2,] = c(inter3[1],inter3[2],mean3)
qtls[3,] = c(inter2[1],inter2[2],mean2)
df = data.frame(qtls)
colnames(df) = c('p(lambda<x)=0.025','p(lambda<x)=0.975','mean')
row.names(df) = c('pior gamma(0.01,0.01)','pior gamma(0.1,0.1)','pior gamma(1,1)')
knitr::kable(df, format = "markdown")
```

The credible interval 95% and the mean are pretty similar for Gamma(0.1,0.1) and Gamma(1,1), Gamma(1,1). The sensitvity for this change of pior is low.

## (4)

```{r}
Y_lst = c(rep(1,12),rep(2,6),rep(10,2),rep(0,30))
a = 0.01
b = 0.01
N = 50
ap =  sum(Y_lst)+a
bp =  (N+b)
x_lst = seq(0,10,by=0.001)
y_lst = dgamma(x_lst,shape=ap,scale=1/bp)
hist(Y_lst,breaks=40,freq=FALSE,ylim = c(0,5),xlim=c(0,10),las=1)
lines(x_lst,y_lst,type='l',col='blue',xlab='lambda',ylab='probability',lwd = 2.5)
inter = qgamma(c(0.025,0.975), shape=ap, scale =1/bp)
I= x_lst>qgamma(.025,shape=ap,scale=1/bp)
x_lstnew = x_lst[I==TRUE]
I= x_lstnew <qgamma(.975,shape=ap,scale=1/bp)
y_lstnew = dgamma(x_lstnew,shape=ap,scale=1/bp)
polygon(c(x_lstnew[I],rev(x_lstnew[I])),c(y_lstnew[I],+ rep(0,sum(I))),col="red",density=30,border=NA)
mean1 =  ap/bp
abline(v = mean1, col="blue", lwd=1.2, lty=2)
```
The posterior distribution dosen't represent the high ASE patient well since it says the pobability denssity is pretty low at 10 side effects and 0 side effects.

## (5)
```{r}
post_samples = rgamma(1000000, shape=ap, scale=1/bp)
pred_samples = sapply(post_samples, rpois, n=1)
hist(pred_samples, breaks = 20, freq = FALSE)
p = sum(pred_samples>1) / length(pred_samples)
print(p)
```

The posterior probablity that has higher side effect than one is 
around 22%. The result is not sensitvity to prior since the posterior possion gamma distribution is not sensitive to the prior.


# Problem 3
From https://en.wikipedia.org/wiki/Gamma_distribution, properties Median calculation
$$100=var=\frac{a}{b^2} => a=100b^2=>b = \frac{\sqrt{a}}{10}$$
$$\mu = \frac{a}{b} = 10 \sqrt{a}$$
$$75= median \approx \mu  \frac{3a-0.8}{3a+0.2} = 10 \sqrt{a} \frac{3a-0.8}{3a+0.2} $$
```{r}
fun <- function (x) 10*(x)^.5*(3*x-0.8)/(3*x+0.2)-75
a <- uniroot(fun, c(0, 100))$root
b = sqrt(a)/10
print(a)
print(b)
```

$$a=56.9139, b= 0.7544132 $$

```{r}
ratt = 1-pgamma(75,a,b)
```
$$P(\lambda>\lambda_{0})= 0.5000023$$